PRIMUS: Galaxy Environment on the Quiescent Fraction Evolution at $z < 0.8$

### 2014-11-21 qfenv_paper/qf_environment.tex
Chang Edits
- Included expected n_env statement in Environment Section
- Adjusted SMF discussion slightly, after SMF changes from new aperture
size.

\title{PRIMUS: Galaxy Environment on the Quiescent Fraction Evolution at $z < 0.8$}
\author{
ChangHoon~Hahn\altaffilmark{1}, 
Michael~Blanton\altaffilmark{1}, 
John Moustakas\altaffilmark{2},
Alison~Coil\altaffilmark{3}, 
Richard~Cool\altaffilmark{4}, 
Daniel Eisenstein\altaffilmark{5},
Ken Wong\altaffilmark{6}, 
Guangtun Zhu\altaffilmark{7}
}
\altaffiltext{1}{Center for Cosmology and Particle Physics, Department of Physics, New York University, 4 Washington Place, New York, NY 10003}
\altaffiltext{2}{Department of Physics and Astronomy, Siena College, 515 Loudon Road, Loudonville, NY 12211}
\altaffiltext{3}{Center for Astrophysics and Space Sciences, Department of Physics, University of California, 9500 Gilman Dr., La Jolla, CA 92093}
\altaffiltext{4}{MMT Observatory, University of Arizona, 1540 E Second Street, Tucson AZ 85721}
\altaffiltext{5}{Harvard-Smithsonian Center for Astrophysics, 60 Garden Street, Cambridge, MA 02138}
\altaffiltext{6}{Steward Observatory, University of Arizona, 933 North Cherry Avenue, Tucson, AZ 85721} 
\altaffiltext{7}{Department of Physics and Astronomy, The Johns Hopkins University, 3400 North Charles Street, Baltimore, MD 21218} 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ABSTRACT
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
We investigate the effects of galaxy environment on the evolution of
the quiescent fraction ($f_{\rm{Q}}$) from $z =0.8 $ to $ 0.0$ using
spectroscopic redshifts and multi-wavelength imaging data from the
PRIsm MUlti-object Survey (PRIMUS) and the Sloan Digitial Sky Survey
(SDSS). Our stellar mass limited galaxy sample consists of $\sim
13,000$ PRIMUS galaxies within $z = 0.2-0.8$ and $\sim 64,000$ SDSS
galaxies within $z = 0.05-0.12$. We classify the galaxies as quiescent
or star-forming based on an evolving specific star formation cut, and
as low or high density environments based on fixed cylindrical
aperture environment measurements on a volume-limited environment
defining population (from PRIMUS and SDSS). For quiescent and
star-forming galaxies in low or high density environments, we examine
the evolution of their stellar mass function (SMF). Then using the
SMFs we compute $f_{\rm{Q}}(\mathcal{M}_{*})$ and quantify its
evolution within our redshift range. We find that the quiescent
fraction is higher at higher masses and in denser environments. The
quiescent fraction rises with cosmic time for all masses and
environments. At a fiducial mass of $10^{10.5}M_\odot$, from $z\sim
0.7$ to $0.1$, the quiescent fraction rises by $15\%$ at the
lowest environments and by $25\%$ at the highest environments we measure.
These results suggest that for a minority of galaxies their cessation
of star formation is due to external influences on
them. However, in the recent Universe a substantial fraction of the
galaxies that cease forming stars do so due to internal processes.
\end{abstract}

@@

\subsection{Galaxy Environment} \label{sec:environment}
We define the environment of a galaxy as the number of neighboring Environment Defining Population galaxies (defined below) within a fixed aperture centered around it. We use a fixed aperture measurement of environment to quantify environment with an aperture sufficiently large to encompass massive halos (\citealt{Muldrew:2012aa, Skibba:2013aa}).
%We use a fixed aperture measurement of environment in order to probe environment on the halo scale, as \cite{Muldrew:2012aa} finds in their comparison of different environment definition using simulations.

@@

In constructing the PRIMUS EDP we use the same PRIMUS data used to select our galaxy sample (described in Section \ref{sec:target}). We again restrict the PRIMUS galaxies to $0.2 < z < 0.8$ and divide them into bins of $\Delta z = 0.2$. Before we consider the cumulative number densities in the redshift bins, we first determine the $M_r$ limit for the highest redshift bin ($0.6-0.8$) by examining the $M_{r}$ distribution with bin size $\Delta M_{r} = 0.25$ and select $M_{r,\rm{lim}}$ near the peak of the distribution where bins with $M_{r} > M_{r,\rm{lim}}$ have fewer galaxies than the bin at $M_{r, \rm{lim}}$. We conservatively choose $M_{r, \rm{lim}}(0.6 < z < 0.8)$ to be $M_{r} = -21.0$. Then for the lower redshift bins, we impose absolute magnitude limits ($M_{r,\rm{lim}}$) such that the cumulative number density, calculated with the galaxy statistical weights, of the bin ordered by $M_{r}$ is equal to the cumulative number density of the highest redshift bin at $M_{r, \rm{lim}}(0.6 < z < 0.8) = -21.0$.

For the SDSS EDP, we do not use the SDSS-{\em GALEX} parent data,
which is limited to the combined angular selection window of the
VAGC and {\em GALEX} (Section \ref{sec:sdss}). Instead, since FUV, NUV
values are not necessary for the EDP, we extend the parent data of the
SDSS EDP to the entire NYU VAGC, including galaxies outside of the
{\em GALEX} window function. Furthermore, we impose a redshift range
of $0.05-0.12$ on the SDSS EDP. This redshift range was determined to
account for the lack of faint galaxies at $z \sim 0.2$ and the lack of
bright galaxies at $z \sim 0.01$ in the VAGC. As with the PRIMUS
redshift bins, we determine the SDSS EDP $M_{r, \rm{lim}}$ by matching
the cumulative number density of the highest redshift bin. For
redshift bins $0.05-0.12$, $0.2-0.4$, $0.4-0.6$, $0.6-0.8$ we get
$M_{r,\rm{lim}} = -20.90$, $-21.06$, $-21.10$ and $-20.98$,
respectively. These absolute magnitude limits are illustrated in
Figure \ref{fig:targetEDP}, where we present the absolute magnitude ($M_{r}$) versus redshift for the galaxy sample (black squares) ad the EDP (red circles). 
The left-most panel corresponds to the sample derived from the SDSS-{\em GALEX} data and the rest correspond to the target sample derived from the PRIMUS data divided in bins with $\Delta z \sim 0.2$. 
Figure \ref{fig:targetEDP} shows clear $M_r$ cutoffs in the
$M_{r}$ distribution versus redshift for the EDP on top
of the target galaxy sample.

@@

Finally, we measure the environment for each galaxy in our galaxy
sample by counting the number of EDP galaxies, $n_{\rm{env}}$ with RA,
Dec, and $z$ within our cylindrical aperture centered around
it. $n_{\rm{env}}$ accounts for the statistical weights of the EDP
galaxies. Once we obtain environment measurements for all the galaxies
in our galaxy sample, we classify galaxies with $n_{\rm{env}} < 0.5$
to be ``low" environment density and galaxies with $n_{\rm{env}} > 3$
to be ``high" environment densities. The high environment cutoff was
selected in order to reduce contamination from galaxies in low
environment densities while maintaining sufficient statistics. In
Section \ref{sec:env_qf_evol} we will also explore higher density
cutoffs for $n_{\rm{env}}$.

@@

While a detailed comparison of the SMFs in each panel for different
epochs is complicated due to the different stellar mass completeness limits, we present some notable trends in each panel. In panel (a), star-forming galaxies in low density environments, we find a significant decrease in the high mass end of the SMF ($\mathcal{M} > 10^{10.75} \mathcal{M}_{\odot}$). At lower masses ($\mathcal{M} < 10^{10.5} \mathcal{M}_{\odot}$) we observe no noticeable trend in the SMF. In panel (b), star-forming galaxies in high density environments, we notice a possible decrease above the knee of the SMF ($\mathcal{M} \sim 10^{10.7} \mathcal{M}_{\odot}$) accompanied by an increase below the knee. For the quiescent population in low density environment, shown in panel (c) of Figure \ref{fig:smf}, we do not notice any clear SMF evolution throughout the mass range. Lastly for the quiescent population in high density environments, panel (d), we find significant increase in $\Phi$ for lower masses but no notable trend at higher masses.

@@

In addition to the evolution over cosmic time, we observe noticeable
trends when we compare the SMFs for star-forming and quiescent
galaxies between the two environments. Comparison of the SMFs in low
versus high density environments reveal a noticeable relation between
mass and density, with SMFs in high density environments having more
massive galaxies. We further confirm this trend w hen we compare the
median mass between the two environments to find that the median mass
for galaxies in high density environments is significantly greater
than in low density environments. The relationship between mass and
environment observed in our SMFs reflects the well-established
mass-density relation and observed mass segregation with environment
in the literature (\citealt{norberg02a, zehavi02a, Blanton:2005ab,
  bundy06a, Scodeggio:2009aa, Bolzonella:2010aa}).

>>>>

\title{PRIMUS: Galaxy Environment on the Quiescent Fraction Evolution at $z < 0.8$}
\author{
ChangHoon~Hahn\altaffilmark{1}, 
Michael~Blanton\altaffilmark{1}, 
John Moustakas\altaffilmark{2},
Alison~Coil\altaffilmark{3}, 
Richard~Cool\altaffilmark{4}, 
Daniel Eisenstein\altaffilmark{5},
Ken Wong\altaffilmark{6}, 
Guangtun Zhu\altaffilmark{7}
}
\altaffiltext{1}{Center for Cosmology and Particle Physics, Department of Physics, New York University, 4 Washington Place, New York, NY 10003}
\altaffiltext{2}{Department of Physics and Astronomy, Siena College, 515 Loudon Road, Loudonville, NY 12211}
\altaffiltext{3}{Center for Astrophysics and Space Sciences, Department of Physics, University of California, 9500 Gilman Dr., La Jolla, CA 92093}
\altaffiltext{4}{MMT Observatory, University of Arizona, 1540 E Second Street, Tucson AZ 85721}
\altaffiltext{5}{Harvard-Smithsonian Center for Astrophysics, 60 Garden Street, Cambridge, MA 02138}
\altaffiltext{6}{Steward Observatory, University of Arizona, 933 North Cherry Avenue, Tucson, AZ 85721} 
\altaffiltext{7}{Department of Physics and Astronomy, The Johns Hopkins University, 3400 North Charles Street, Baltimore, MD 21218} 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ABSTRACT
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
We investigate the effects of galaxy environment on the evolution of
the quiescent fraction ($f_{\rm{Q}}$) from $z =0.8 $ to $ 0.0$ using
spectroscopic redshifts and multi-wavelength imaging data from the
PRIsm MUlti-object Survey (PRIMUS) and the Sloan Digitial Sky Survey
(SDSS). Our stellar mass limited galaxy sample consists of $\sim
13,000$ PRIMUS galaxies within $z = 0.2-0.8$ and $\sim 64,000$ SDSS
galaxies within $z = 0.05-0.12$. We classify the galaxies as quiescent
or star-forming based on an evolving specific star formation cut, and
as low or high density environments based on fixed cylindrical
aperture environment measurements on a volume-limited environment
defining population. For quiescent and
star-forming galaxies in low or high density environments, we examine
the evolution of their stellar mass function (SMF). Then using the
SMFs we compute $f_{\rm{Q}}(\mathcal{M}_{*})$ and quantify its
evolution within our redshift range. We find that the quiescent
fraction is higher at higher masses and in denser environments. The
quiescent fraction rises with cosmic time for all masses and
environments. At a fiducial mass of $10^{10.5}M_\odot$, from $z\sim
0.7$ to $0.1$, the quiescent fraction rises by $15\%$ at the
lowest environments and by $25\%$ at the highest environments we measure.
These results suggest that for a minority of galaxies their cessation
of star formation is due to external influences on
them. However, in the recent Universe a substantial fraction of the
galaxies that cease forming stars do so due to internal processes.
\end{abstract}

@@

\subsection{Galaxy Environment} \label{sec:environment}
We define the environment of a galaxy as the number of neighboring Environment Defining Population galaxies (defined below) within a fixed aperture centered around it. We use fixed aperture measurements in order to quantify galaxy environment with an aperture sufficiently large to encompass massive halos (\citealt{Muldrew:2012aa, Skibba:2013aa}).
%We use a fixed aperture measurement of environment in order to probe environment on the halo scale, as \cite{Muldrew:2012aa} finds in their comparison of different environment definition using simulations.

@@

In constructing the PRIMUS EDP we use the same PRIMUS data used to select our galaxy sample (described in Section \ref{sec:target}). We again restrict the PRIMUS galaxies to $0.2 < z < 0.8$ and divide them into bins of $\Delta z = 0.2$. Before we consider the cumulative number densities in the redshift bins, we first determine the $M_r$ limit for the highest redshift bin ($0.6-0.8$) by examining the $M_{r}$ distribution with bin size $\Delta M_{r} = 0.25$ and select $M_{r,\rm{lim}}$ near the peak of the distribution where bins with $M_{r} > M_{r,\rm{lim}}$ have fewer galaxies than the bin at $M_{r, \rm{lim}}$. We conservatively choose $M_{r, \rm{lim}}(0.6 < z < 0.8)$ to be $M_{r} = -21.0$. Then for the lower redshift bins, we impose absolute magnitude limits ($M_{r,\rm{lim}}$) such that the cumulative number density, calculated with the galaxy statistical weights, of the bin ordered by $M_{r}$ is equal to the cumulative number density of the highest redshift bin at $M_{r, \rm{lim}}(0.6 < z < 0.8) = -20.97$.

For the SDSS EDP, we do not use the SDSS-{\em GALEX} parent data,
which is limited to the combined angular selection window of the
VAGC and {\em GALEX} (Section \ref{sec:sdss}). Instead, since FUV, NUV
values are not necessary for the EDP, we extend the parent data of the
SDSS EDP to the entire NYU VAGC, including galaxies outside of the
{\em GALEX} window function. Furthermore, we impose a redshift range
of $0.05-0.12$ on the SDSS EDP. This redshift range was determined to
account for the lack of faint galaxies at $z \sim 0.2$ and the lack of
bright galaxies at $z \sim 0.01$ in the VAGC. As with the PRIMUS
redshift bins, we determine the SDSS EDP $M_{r, \rm{lim}}$ by matching
the cumulative number density of the highest redshift bin. For
redshift bins $z = 0.05-0.12$, $0.2-0.4$, $0.4-0.6$, $0.6-0.8$ we get
$M_{r,\rm{lim}} = -20.95$, $-21.03$, $-20.98$ and $-20.97$,
respectively. These absolute magnitude limits are illustrated in
Figure \ref{fig:targetEDP}, where we present the absolute magnitude ($M_{r}$) versus redshift for the galaxy sample (black squares) ad the EDP (red circles). 
The left-most panel corresponds to the sample derived from the SDSS-{\em GALEX} data and the rest correspond to the target sample derived from the PRIMUS data divided in bins with $\Delta z \sim 0.2$. 
Figure \ref{fig:targetEDP} shows clear $M_r$ cutoffs in the
$M_{r}$ distribution versus redshift for the EDP on top
of the target galaxy sample.

@@

Finally, we measure the environment for each galaxy in our galaxy
sample by counting the number of EDP galaxies, $n_{\rm{env}}$ with RA,
Dec, and $z$ within our cylindrical aperture centered around
it. $n_{\rm{env}}$ accounts for the statistical weights of the EDP
galaxies. 
For our galaxy sample, the expected $n_{\rm{env}}$ given the constant number density in 
each of our EDP redshift bin and volume of our cylindrical aperture is $< n_{\rm{env}} > = 1.3$. 
Once we obtain environment measurements for all the galaxies
in our galaxy sample, we classify galaxies with $n_{\rm{env}} < 0.5$
to be ``low" environment density and galaxies with $n_{\rm{env}} > 3$
to be ``high" environment densities. The high environment cutoff was
selected in order to reduce contamination from galaxies in low
environment densities while maintaining sufficient statistics. In
Section \ref{sec:env_qf_evol} we will also explore higher density
cutoffs for $n_{\rm{env}}$.

@@

While a detailed comparison of the SMFs in each panel for different
epochs is complicated due to the different stellar mass completeness limits, we present some notable trends in each panel. In panel (a), star-forming galaxies in low density environments, we find a significant decrease in the high mass end of the SMF ($\mathcal{M} > 10^{10.75} \mathcal{M}_{\odot}$). At lower masses ($\mathcal{M} < 10^{10.5} \mathcal{M}_{\odot}$) we observe no noticeable trend in the SMF. In panel (b), star-forming galaxies in high density environments, we do not observe any clear trends above the knee of the SMF ($\mathcal{M} \sim 10^{10.7} \mathcal{M}_{\odot}$) but an increase in SMF below the knee. For the quiescent population in low density environment, shown in panel (c) of Figure \ref{fig:smf}, we observe a potential decrease at high masses ($\mathcal{M} > 10^{10.7} \mathcal{M}_{\odot}$). Lastly for the quiescent population in high density environments, panel (d), we find significant increase in $\Phi$ for lower masses but no notable trend at higher masses. 
%do not notice any clear SMF evolution throughout the mass range.

@@

In addition to the evolution over cosmic time, we observe noticeable
trends when we compare the SMFs for star-forming and quiescent
galaxies between the two environments. Comparison of the SMFs in low
versus high density environments reveal a noticeable relation between
mass and density, with SMFs in high density environments having more
massive galaxies, especially evident in our lowest redshift bin. 
We further confirm this trend when we compare the
%massive galaxies. We further confirm this trend w hen we compare the
median mass between the two environments to find that the median mass
for galaxies in high density environments is significantly greater
than in low density environments. The relationship between mass and
environment observed in our SMFs reflects the well-established
mass-density relation and observed mass segregation with environment
in the literature (\citealt{norberg02a, zehavi02a, Blanton:2005ab,
  bundy06a, Scodeggio:2009aa, Bolzonella:2010aa}).
